---
title: Report
date: '`r format(Sys.time(), "%Y-%m-%d-%H-%M")`'
output:
  workflowr::wflow_html:
    toc: yes
    toc_float: yes
    toc_depth: 3
    code_folding: hide
    theme: cosmo
    highlight: textmate
    df_print: paged
workflowr:
  suppress_report: TRUE
params:
  bin_name: "10_micron"
  family: "constant"
  l: "20"
  prefix: ""
  nn: ""
---

# Packages

```{r}
suppressPackageStartupMessages({
  library(mistyR)
  library(tidyverse)
  library(factoextra)
  library(png)
  library(here)
  library(igraph)
})
```

# Configs

```{r}
here::i_am("scripts/report.Rmd")
bin_name <- params$bin_name
family <- params$family
l <- params$l
prefix <- params$prefix
nn <- params$nn
config_string <- paste0(bin_name, "_", family, "_", l, "_", prefix, "_", nn)
config_string
```

```{r}
source(here::here("config.R"))
source(here::here("scripts", "utils.R"))
```

# Load Data

```{r}
### load processed data ###
file_data <- readRDS(here::here("processed_data", "file_data.RDS"))
coords <- readRDS(here::here("processed_data", paste0("coords_", bin_name, ".RDS")))
cluster_counts <- readRDS(here::here("processed_data", paste0("cluster_counts_", bin_name, ".RDS")))
all(names(coords) == names(cluster_counts)) # check
img_names <- names(cluster_counts)

# Stratify clusters based on whether they are specific or unspecific (see Mail by Malte).
stopifnot(all(sort(c(unspecific_clusters, specific_clusters)) == sort(all_clusters)))

# Prepare indexing by character. The clusters are names from 0 to 46 which is convenient in Python (0-based indexing), 
# but annoying in R (1-based indexing). So I will just index by character.
for (img_name in names(cluster_counts)) {
  # take care of cluster counts
  mtx <- cluster_counts[[img_name]]
  colnames(mtx) <- all_clusters
  cluster_counts[[img_name]] <- mtx
  
  # take care of coordinates
  c_mtx <- coords[[img_name]]
  colnames(c_mtx) <- c("x", "y")
  coords[[img_name]] <- c_mtx
}

in_file <- paste0(processed_data_dir, .Platform$file.sep, "all_views_", config_string, ".RDS")
print(paste0("Reading views from: ", in_file))
all_views <- readRDS(in_file)
```

# Results Global

```{r}
results <- readRDS(here::here(misty_dir, paste0( "global_", config_string, ".RDS")))
name_paraview <- grep("para.+", unique(results$importances$view), value=TRUE)
name_paraview_long <- grep("para.+", names(all_views[[1]]), value=TRUE)
rm(all_views); gc()

results$importances <- results$importances %>%
  dplyr::mutate(Predictor = str_remove(Predictor, "^p_"))

# first compute the mean importances per patient, then aggregate
results$importances.aggregated <- results$importances %>%
  dplyr::mutate(sample=basename(sample)) %>%
  dplyr::left_join(file_data %>% dplyr::select(filename, unique_patient_id),
                   by=c("sample" = "filename")) %>%
  dplyr::group_by(view, Predictor, Target, unique_patient_id) %>%
  dplyr::summarise(Importance = mean(Importance), .groups = "drop_last") %>%
  dplyr::summarise(Importance = mean(Importance), nsamples = n(), .groups = "drop")
```

## Performance

```{r}
results %>%
  plot_improvement_stats("gain.R2")
```

## View Contributions

```{r}
results %>% plot_view_contributions()
```

## Importances

```{r fig.width=8, fig.height=8}
results %>% plot_interaction_heatmap(view = "intra", cutoff = 0, clean=TRUE)
```

```{r fig.width=8, fig.height=8}
results %>% plot_interaction_heatmap(view = name_paraview, cutoff = 0, clean=TRUE)
```

## Signatures

### Performance

```{r}
perf_sign <- mistyR::extract_signature(results, "performance")
perf_sign_pca <- stats::prcomp(perf_sign %>% select(matches("gain.R2")))
```

No clusters here.

```{r}
ggplot() +
  geom_point(aes(x=perf_sign_pca$x[,1], y=perf_sign_pca$x[,2]))
```

### Importances

```{r}
imp_sign <- mistyR::extract_signature(results, "importance")
imp_sign_pca <- stats::prcomp(imp_sign %>% dplyr::select(-sample))
```

We see two clusters in the first principal components.

```{r}
tibble(x=imp_sign_pca$x[,1], y=imp_sign_pca$x[,2]) %>%
  ggplot() +
  geom_point(aes(x=x, y=y))
```

What drives the seperation?

```{r}
fviz_pca_var(imp_sign_pca, col.var="contrib",
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE, select.var = list("contrib"=20))
```

What could explain the separation?

Looking at the sample metadata we see that we can clearly distinguish Diabetic-Neophropathy vs CKD/Non-CKD. So this should be reflected in the importance contrasts.

```{r fig.width=10, fig.height=8}
purrr::map(c("group_new", "egfr_mdrd", "ckd_stage", "age", "ischemic_time", "sex_m0f1"), function(v) {
  v <- sym(v)
  tibble(x=imp_sign_pca$x[,1], y=imp_sign_pca$x[,2],
         sample=basename(imp_sign$sample)) %>%
    dplyr::left_join(file_data, by=c("sample"="filename")) %>%
    dplyr::mutate(sex_m0f1 = as.factor(sex_m0f1)) %>%
    ggplot() +
    geom_point(aes(x=x, y=y, color=!!v), size=1)
}) %>%
  cowplot::plot_grid(plotlist=., ncol=2)
```

Let's look at the first 3 principal components.

```{r fig.width=10, fig.height=4}
p1 <- tibble(x=imp_sign_pca$x[,1], y=imp_sign_pca$x[,2],
       sample=basename(imp_sign$sample)) %>%
  dplyr::left_join(file_data, by=c("sample"="filename")) %>%
  ggplot() +
  geom_point(aes(x=x, y=y, color=group_new), size=1, show.legend = F) +
  labs(x="PCA 1", y="PCA 2", color="condition")

p2 <- tibble(x=imp_sign_pca$x[,2], y=imp_sign_pca$x[,3],
       sample=basename(imp_sign$sample)) %>%
  dplyr::left_join(file_data, by=c("sample"="filename")) %>%
  ggplot() +
  geom_point(aes(x=x, y=y, color=group_new), size=1) +
  labs(x="PCA 2", y="PCA 3", color="condition")

cowplot::plot_grid(p1, p2, rel_widths=c(1, 1.5))
```

Note that the abundance of clusters per image can already explain these differences.

```{r}
cluster_count_per_image <- purrr::imap_dfr(cluster_counts, function(mtx, name) {
  c("image" = name, colSums(mtx))
}) %>%
  dplyr::mutate(across(!image, ~ as.numeric(.x)))

tibble(x=imp_sign_pca$x[,1], y=imp_sign_pca$x[,2],
         sample=basename(imp_sign$sample)) %>%
  dplyr::left_join(cluster_count_per_image, by=c("sample"="image")) %>%
  dplyr::select(x, y, sample, `19`, `2`, `13`, `14`, `41`, `21`) %>%
  tidyr::pivot_longer(cols=!(c(x, y, sample))) %>%
  dplyr::mutate(value=log10(value)) %>%
  ggplot(aes(x=x, y=y, color=value)) +
  geom_point() +
  facet_wrap(~name) +
  scale_color_viridis_c() +
  labs(color="Log10 Counts")
```

Here are the PCA results, though the seperation when running MISTy looks somewhat better.

```{r fig.width=10, fig.height=8}
abund_pca <- stats::prcomp(cluster_count_per_image %>% 
                             dplyr::select(-image))

purrr::map(c("group_new", "egfr_mdrd", "ckd_stage", "age", "ischemic_time", "sex_m0f1"), function(v) {
  v <- sym(v)
  tibble(x=abund_pca$x[,1], y=abund_pca$x[,2],
         sample=cluster_count_per_image$image) %>%
    dplyr::left_join(file_data, by=c("sample"="filename")) %>%
    dplyr::mutate(sex_m0f1 = as.factor(sex_m0f1)) %>%
    ggplot() +
    geom_point(aes(x=x, y=y, color=!!v), size=1)
}) %>%
  cowplot::plot_grid(plotlist=., ncol=2)
```

```{r}
rm(cluster_counts, results)
gc()
```

# Results per Group

```{r}
results.groups <- readRDS(here::here(misty_dir, paste0("DKD_non-DKD_", 
                                                       config_string, ".RDS")))

for (group in names(results.groups)) {
  results.groups[[group]]$importances <- results.groups[[group]]$importances %>%
    dplyr::mutate(Predictor = str_remove(Predictor, "^p_"))
  
  # first compute the mean importances per patient, then aggregate
  results.groups[[group]]$importances.aggregated <- 
    results.groups[[group]]$importances %>%
      dplyr::mutate(sample=basename(sample)) %>%
      dplyr::left_join(file_data %>% dplyr::select(filename, unique_patient_id),
                       by=c("sample" = "filename")) %>%
      dplyr::group_by(view, Predictor, Target, unique_patient_id) %>%
      dplyr::summarise(Importance = mean(Importance), .groups = "drop_last") %>%
      dplyr::summarise(Importance = mean(Importance), nsamples = n(), .groups = "drop")
}
```

```{r}
cluster_props_per_cond <- cluster_count_per_image %>%
  dplyr::select(image, all_of(specific_clusters)) %>%
  dplyr::mutate(unspec = rowSums(cluster_count_per_image[, unspecific_clusters])) %>%
  tidyr::pivot_longer(cols=!image, names_to="cluster", values_to="count") %>%
  dplyr::left_join(file_data %>% dplyr::select(filename, group_new), by=c("image"="filename")) %>%
  dplyr::group_by(group_new, cluster) %>%
  dplyr::summarise(count = mean(count), .groups = "drop_last") %>%
  dplyr::mutate(fraction = count / sum(count) * 100, count=NULL) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(cluster = paste0("c_", cluster))
```

## Performance

```{r fig.width=16, fig.height=4}
purrr::map(names(results.groups), function(cond) {
  results.groups[[cond]]$improvements %>%
  dplyr::filter(measure=="intra.R2") %>%
  dplyr::group_by(target) %>%
  dplyr::summarise(mean=mean(value), sd=sd(value), .groups="drop") %>%
  ggplot(aes(x=reorder(target, -mean))) +
  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd)) +
  geom_point(aes(y=mean)) +
  theme(axis.text.x = element_text(angle=45, hjust=1), 
        plot.title = element_text(hjust=0.5)) +
  labs(x="Clusters", y="Intra R2", title=cond)
}) %>%
  cowplot::plot_grid(plotlist=., ncol=2)

purrr::map(names(results.groups), function(cond) {
  results.groups[[cond]]$improvements %>%
  dplyr::filter(measure=="multi.R2") %>%
  dplyr::group_by(target) %>%
  dplyr::summarise(mean=mean(value), sd=sd(value), .groups="drop") %>%
  ggplot(aes(x=reorder(target, -mean))) +
  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd)) +
  geom_point(aes(y=mean)) +
  theme(axis.text.x = element_text(angle=45, hjust=1), 
        plot.title = element_text(hjust=0.5)) +
  labs(x="Clusters", y="Multi R2", title=cond)
}) %>%
  cowplot::plot_grid(plotlist=., ncol=2)

purrr::map(names(results.groups), function(cond) {
  results.groups[[cond]]$improvements %>%
  dplyr::filter(measure=="gain.R2") %>%
  dplyr::group_by(target) %>%
  dplyr::summarise(mean=mean(value), sd=sd(value), .groups="drop") %>%
  ggplot(aes(x=reorder(target, -mean))) +
  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd)) +
  geom_point(aes(y=mean)) +
  theme(axis.text.x = element_text(angle=45, hjust=1), 
        plot.title = element_text(hjust=0.5)) +
  labs(x="Clusters", y="Gain R2", title=cond)
}) %>%
  cowplot::plot_grid(plotlist=., ncol=2)
```

How much is the performance related to the abundance?

```{r fig.width=12, fig.height=4}
purrr::map_dfr(names(results.groups), function(cond) {
  results.groups[[cond]]$improvements %>%
    dplyr::filter(measure=="multi.R2") %>%
    dplyr::group_by(target) %>%
    dplyr::summarise(mean=mean(value), sd=sd(value), .groups="drop") %>%
    dplyr::mutate(group_new = cond)
}) %>%
  dplyr::left_join(cluster_props_per_cond, by=c("group_new"="group_new", "target"="cluster")) %>%
  ggplot(aes(x=fraction, y=mean)) +
  geom_point() +
  facet_wrap(~group_new) +
  labs(x="Cluster Fraction [%]", y="Multi R2 [%]")

# without unspecific
purrr::map_dfr(names(results.groups), function(cond) {
  results.groups[[cond]]$improvements %>%
    dplyr::filter(measure=="multi.R2") %>%
    dplyr::group_by(target) %>%
    dplyr::summarise(mean=mean(value), sd=sd(value), .groups="drop") %>%
    dplyr::mutate(group_new = cond)
}) %>%
  dplyr::left_join(cluster_props_per_cond, by=c("group_new"="group_new", "target"="cluster")) %>%
  dplyr::filter(target != "c_unspec") %>%
  ggplot(aes(x=fraction, y=mean)) +
  geom_point() +
  facet_wrap(~group_new) +
  labs(x="Mean Cluster Fraction [%]", y="Multi R2 [%]", title="Without the Unspecific Cluster")
```

## Importances

### Control

```{r}
group_oi <- "Control"
```

```{r}
results.groups[[group_oi]] %>% plot_interaction_heatmap(view = "intra", cutoff = 0, clean=TRUE)
results.groups[[group_oi]] %>% plot_interaction_heatmap(view = name_paraview, cutoff = 0, clean=TRUE)
```

```{r}
concat_views_oi <- readRDS(paste0(processed_data_dir, .Platform$file.sep, "concat_views_", config_string, "_", group_oi, ".RDS"))
```

```{r fig.width=8, fig.height=8}
conf_list <- list(view.short="intra", view.long="intraview", file.name="intra",
                  cutoff=1.2, resolution=0.3)
out <- interaction_communities_info(misty.results=results.groups[[group_oi]], 
                                    concat.views=concat_views_oi, 
                                    view.short=conf_list$view.short, 
                                    view.long=conf_list$view.long,
                                    cutoff=conf_list$cutoff, res=conf_list$resolution)
color_func <- circlize::colorRamp2(c(-1, 0, 1), c("blue", "white", "red"))
plot.igraph(out, 
            layout=layout_with_fr(out),
            edge.curved=0.75,
            vertex.color=igraph::vertex_attr(out)$community,
            edge.color=color_func(igraph::edge_attr(out)$cor),
            edge.width=scales::rescale(igraph::edge_attr(out)$weight, to=c(1,10)))
```

```{r fig.width=8, fig.height=8}
conf_list <- list(view.short=name_paraview, view.long=name_paraview_long, file.name="para",
                  cutoff=0.8, resolution=1)
out <- interaction_communities_info(misty.results=results.groups[[group_oi]], 
                                    concat.views=concat_views_oi, 
                                    view.short=conf_list$view.short, 
                                    view.long=conf_list$view.long,
                                    cutoff=conf_list$cutoff, res=conf_list$resolution)

color_func <- circlize::colorRamp2(c(-1, 0, 1), c("blue", "white", "red"))
plot.igraph(out, 
            layout=layout_with_fr(out),
            edge.curved=0.75,
            vertex.color=igraph::vertex_attr(out)$community,
            edge.color=color_func(igraph::edge_attr(out)$cor),
            edge.width=scales::rescale(igraph::edge_attr(out)$weight, to=c(1,10)))
```

### Diabetic-Nephropathy

```{r}
group_oi <- "Diabetic-Nephropathy"
```

```{r}
results.groups[[group_oi]] %>% plot_interaction_heatmap(view = "intra", cutoff = 1, clean=TRUE)
results.groups[[group_oi]] %>% plot_interaction_heatmap(view = name_paraview, cutoff = 1, clean=TRUE)
```

```{r}
concat_views_oi <- readRDS(paste0(processed_data_dir, .Platform$file.sep, "concat_views_", config_string, "_", group_oi, ".RDS"))
```

```{r fig.width=8, fig.height=8}
conf_list <- list(view.short="intra", view.long="intraview", file.name="intra",
                  cutoff=1.2, resolution=0.3)
out <- interaction_communities_info(misty.results=results.groups[[group_oi]], 
                                    concat.views=concat_views_oi, 
                                    view.short=conf_list$view.short, 
                                    view.long=conf_list$view.long,
                                    cutoff=conf_list$cutoff, res=conf_list$resolution)
color_func <- circlize::colorRamp2(c(-1, 0, 1), c("blue", "white", "red"))
plot.igraph(out, 
            layout=layout_with_fr(out),
            edge.curved=0.75,
            vertex.color=igraph::vertex_attr(out)$community,
            edge.color=color_func(igraph::edge_attr(out)$cor),
            edge.width=scales::rescale(igraph::edge_attr(out)$weight, to=c(1,10)))
```

```{r fig.width=8, fig.height=8}
conf_list <- list(view.short=name_paraview, view.long=name_paraview_long, file.name="para",
                  cutoff=0.8, resolution=1)
out <- interaction_communities_info(misty.results=results.groups[[group_oi]], 
                                    concat.views=concat_views_oi, 
                                    view.short=conf_list$view.short, 
                                    view.long=conf_list$view.long,
                                    cutoff=conf_list$cutoff, res=conf_list$resolution)
color_func <- circlize::colorRamp2(c(-1, 1), c("blue", "red"))
plot.igraph(out, 
            layout=layout_with_fr(out),
            edge.curved=0.75,
            vertex.color=igraph::vertex_attr(out)$community,
            edge.color=color_func(igraph::edge_attr(out)$cor),
            edge.width=scales::rescale(igraph::edge_attr(out)$weight, to=c(1,10)))
```

## Importance Contrasts: Control vs. Diabetic-Nephropathy

```{r}
plot_contrast_results(misty.results.from=results.groups$Control,
                      misty.results.to=results.groups$`Diabetic-Nephropathy`,
                      views=c("intra", name_paraview))
```







